---
title: "Exploratory Data Analysis"
author: "Spiro Stilianoudakis"
date: "9/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r}
library(openxlsx)
library(dplyr)
library(tableone)
```

# Setting working directory

```{r}
setwd("~/Biostatistical_Consulting_Lab/R168/Data")
```


# Patients with behavioral event

```{r}
setwd("~/Biostatistical_Consulting_Lab/R168/Data")
pwbe <- read.csv("keppra_data_with_behavior_v3.csv", 
                header = TRUE,
                na.strings = "N/A",
                stringsAsFactors = FALSE,
                nrows = 453)
                #colClasses = c(rep("numeric",2),
                #               rep("character",3),
                #               rep("numeric",1),
                #               rep("character",1),
                #               rep("numeric",3),
                #               rep("character",3),
                #               rep("numeric",1),
                #               rep("character",1),
                #               rep("numeric",1),
                #               rep("character",2)))

dim(pwbe)
#453  30

table(unlist(lapply(pwbe,class)))
#character   integer   numeric 
#       17         6         7 
```

# Patients without behavioral event

```{r}
setwd("~/Biostatistical_Consulting_Lab/R168/Data")
pwobe <- read.csv("keppra_data_without_behavior_v3.csv", 
                header = TRUE,
                na.strings='N/A',
                stringsAsFactors = FALSE,
                nrows = 512)
                #colClasses = c(rep("numeric",2),
                #               rep("character",3),
                #               rep("numeric",1),
                #               rep("character",1),
                #               rep("numeric",3)))

dim(pwobe)
#512  22

table(unlist(lapply(pwobe,class)))
#character   integer   numeric 
#       11         6         5 
```

#Combine with and without behavioral event with only demographics

```{r}
keppra_demo <- rbind.data.frame(pwbe[,names(pwbe) %in% intersect(names(pwbe), names(pwobe))],
                                pwobe[,names(pwobe) %in% intersect(names(pwobe), names(pwbe))])

dim(keppra_demo)
#965  22

keppra_demo$BehavioralEvent <- c(rep("Yes", dim(pwbe)[1]),
                                 rep("No", dim(pwobe)[1]))
table(keppra_demo$BehavioralEvent)
# No Yes 
#512 453

names(keppra_demo) <- c("ID",
                        "Age",
                        "Ethnicity",
                        "Gender",
                        "Diagnosis",
                        "GCS",
                        "Discharge.Disposition",
                        "Length_of_Stay_Hospital",
                        "Length_of_Stay_ICU",
                        "Total_Keppra_Doses",
                        "Keppra_Duration",
                        "Rass_less_than_3",
                        "Median_Keppra_Dose",
                        "Received_Benzo",
                        "CIWA_Orders",
                        "Pain_Scores",
                        "Received_IV_opioid",
                        "Concurrent_AED",
                        "AED_Received_while_on_Keppra",
                        "Times_RASS_less_neg3",
                        "Times_RASS_evaluated",
                        "Percent_RASS_less_neg3",
                        "Behavioral_Event")
```

# Checking missingness 

```{r}
#missingness
lapply(keppra_demo, function(x)table(is.na(x))) 
##missingness from the "AED Received while on Keppra" variable
##missingness from the Times_RASS_less_neg3 variable
##missingness from the Times_RASS_evaluated variable
##missingness from the Percent_RASS_less_neg3 variable

#summaries
lapply(keppra_demo, function(x)if(class(x)=="numeric" | class(x)=="integer"){summary(x)}else(table(x)))
##ethnicity needs recoding
##rass needs recoding
##median keppra dose needs recoding
##AED_Received_while_on_Keppra needs recoding
##pain score needs recoding

```

# Recoding variables

## Ethnicity

```{r}
##black, white, other, unknown
keppra_demo$Ethnicity[which(keppra_demo$Ethnicity %in% c("American Indian-Alaskan", "Asian", "Multiple", "Native Hawaii/Other Pac Island", "Other"))] <- "Other"
keppra_demo$Ethnicity[which(keppra_demo$Ethnicity %in% c("N/A-Outreach Use Only", "Unknown-Unable to Communicate", "Unknown-Pt Refusal"))] <- "Unknown"
table(keppra_demo$Ethnicity)

keppra_demo$Ethnicity <- factor(keppra_demo$Ethnicity, levels=c("White", "Black or African American", "Other", "Unknown"))

```

## Rass

```{r}
keppra_demo$Rass_less_than_3[which(keppra_demo$Rass_less_than_3=="yes")] <- "Yes" 
table(keppra_demo$Rass_less_than_3)
```

## Median keppra dose needs recoding

```{r}
keppra_demo$Median_Keppra_Dose[which(keppra_demo$Median_Keppra_Dose=="5/14/1901")] <- "Missing"

keppra_demo$Median_Keppra_Dose <- factor(keppra_demo$Median_Keppra_Dose, levels = c("250", "500", "750", "1000", "1250", "1500", "1750", "2000", "3000", "Missing"))

table(keppra_demo$Median_Keppra_Dose)

```

## AED Received while on Keppra 

```{r}
keppra_demo$AED_Received_while_on_Keppra <- ifelse(is.na(keppra_demo$AED_Received_while_on_Keppra), "No", "Yes")
table(keppra_demo$AED_Received_while_on_Keppra)
```

## Pain score

```{r}
table(keppra_demo$Pain_Scores, exclude = "ifany")
keppra_demo$Pain_Scores[which(keppra_demo$Pain_Scores=="10-Apr")] <- "4-10"
table(keppra_demo$Pain_Scores, exclude = "ifany")
```


#Creating table of demographics

```{r}
#create log of variables
keppra_demo$log_Length_of_Stay_Hospital <- log10(keppra_demo$Length_of_Stay_Hospital)
keppra_demo$log_Length_of_Stay_ICU <- log10(keppra_demo$Length_of_Stay_ICU)
keppra_demo$log_Total_Keppra_Doses <- log10(keppra_demo$Total_Keppra_Doses)
keppra_demo$log_Keppra_Duration <- log10(keppra_demo$Keppra_Duration+1)

tableOne <- CreateTableOne(vars = c("Age",
                        "Ethnicity",
                        "Gender",
                        "Diagnosis",
                        "GCS",
                        "Discharge.Disposition",
                        "Length_of_Stay_Hospital",
                        "Length_of_Stay_ICU",
                        "Total_Keppra_Doses",
                        "Keppra_Duration",
                        "log_Length_of_Stay_Hospital",
                        "log_Length_of_Stay_ICU",
                        "log_Total_Keppra_Doses",
                        "log_Keppra_Duration",
                        "Rass_less_than_3",
                        "Median_Keppra_Dose",
                        "Received_Benzo",
                        "CIWA_Orders",
                        "Pain_Scores",
                        "Received_IV_opioid",
                        "Concurrent_AED",
                        "AED_Received_while_on_Keppra",
                        "Times_RASS_less_neg3",
                        "Times_RASS_evaluated",
                        "Percent_RASS_less_neg3"
                        ),
               strata = "Behavioral_Event",
               data = keppra_demo)


```

# Table of only surviving patients (icu length of stay, hospital length of stay, keppra dose, and keppra duration)

```{r}
keppra_demo_survived <- keppra_demo[-which(keppra_demo$Discharge.Disposition=="Expired"),]

dim(keppra_demo_survived) #805  27

tableOne_survived <- CreateTableOne(vars = c("Age",
                        "Ethnicity",
                        "Gender",
                        "Diagnosis",
                        "GCS",
                        "Discharge.Disposition",
                        "Length_of_Stay_Hospital",
                        "Length_of_Stay_ICU",
                        "Total_Keppra_Doses",
                        "Keppra_Duration",
                        "log_Length_of_Stay_Hospital",
                        "log_Length_of_Stay_ICU",
                        "log_Total_Keppra_Doses",
                        "log_Keppra_Duration",
                        "Rass_less_than_3",
                        "Median_Keppra_Dose",
                        "Received_Benzo",
                        "CIWA_Orders",
                        "Pain_Scores",
                        "Received_IV_opioid",
                        "Concurrent_AED",
                        "AED_Received_while_on_Keppra",
                        "Times_RASS_less_neg3",
                        "Times_RASS_evaluated",
                        "Percent_RASS_less_neg3"
                        ),
               strata = "Behavioral_Event",
               data = keppra_demo_survived)

```

