---
title: "Incidences for patients with behavioral event"
author: "Spiro Stilianoudakis"
date: "9/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r}
library(openxlsx)
library(dplyr)
library(tableone)
library(survival)
library(ggplot2)
library(ggfortify)
library(survminer)
```

# Patients with behavioral event

```{r}
pwbe <- read.csv("Y:/Shares/Biostatistics/Projects/BCL/Research Projects/Open/BCL - 2018 - R168 (Holton-Burke - Nerology)/1 - Data/keppra_data_with_behavior.csv", 
                header = TRUE,na.strings='N/A',
                colClasses = c(rep("numeric",2),
                               rep("character",3),
                               rep("numeric",1),
                               rep("character",1),
                               rep("numeric",3),
                               rep("character",3),
                               rep("numeric",1),
                               rep("character",1),
                               rep("numeric",1),
                               rep("character",2)))

dim(pwbe)
#453  18

table(unlist(lapply(pwbe,class)))
#character   numeric 
#       10         8 
```

# Incidence

## Delirium Diagnosis

```{r}
table(pwbe$Delirium.diagnosis, exclude = "ifany")
# No Yes 
#305 148

pwbe %>% 
    group_by(Delirium.diagnosis) %>% 
    summarize(count = n()) %>% 
    mutate(percent = count/sum(count)) %>% 
    ggplot(aes(x=Delirium.diagnosis, y=count, fill=Delirium.diagnosis)) +
    geom_col() +
    geom_text(aes(label = paste0(round(100 * percent, 1), "%")), vjust = -.5, size=5)+
    xlab("Delirium Diagnosis") + 
    ylab("Count")+
    guides(fill=FALSE)+
    theme_bw()+
    theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20))
```

## Antipsychotic administration

```{r}
table(pwbe$Antipsychotic.administration.while.on.Keppra, exclude = "ifany")
# No Yes 
#262 191

pwbe %>% 
    group_by(Antipsychotic.administration.while.on.Keppra) %>% 
    summarize(count = n()) %>% 
    mutate(percent = count/sum(count)) %>% 
    ggplot(aes(x=Antipsychotic.administration.while.on.Keppra, y=count, fill=Antipsychotic.administration.while.on.Keppra)) +
    geom_col() +
    geom_text(aes(label = paste0(round(100 * percent, 1), "%")), vjust = -.5, size=5)+
    xlab("Received Antipsychotic") + 
    ylab("Count")+
    guides(fill=FALSE)+
    theme_bw()+
    theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20))
```

## Positive CAM-ICU

```{r}
table(pwbe$Positive.CAM.ICU.while.on.Keppra, exclude = "ifany")
# No Yes 
#340 113

pwbe %>% 
    group_by(Positive.CAM.ICU.while.on.Keppra) %>% 
    summarize(count = n()) %>% 
    mutate(percent = count/sum(count)) %>% 
    ggplot(aes(x=Positive.CAM.ICU.while.on.Keppra, y=count, fill=Positive.CAM.ICU.while.on.Keppra)) +
    geom_col() +
    geom_text(aes(label = paste0(round(100 * percent, 1), "%")), vjust = -.5, size=5)+
    xlab("Positive CAM-ICU") + 
    ylab("Count")+
    guides(fill=FALSE)+
    theme_bw()+
    theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20))
```

## Adverse event disgnosis code

```{r}
table(pwbe$Adverse.event.diagnosis.code, exclude = "ifany")
# No Yes 
#193 260 

pwbe %>% 
    group_by(Adverse.event.diagnosis.code) %>% 
    summarize(count = n()) %>% 
    mutate(percent = count/sum(count)) %>% 
    ggplot(aes(x=Adverse.event.diagnosis.code, y=count, fill=Adverse.event.diagnosis.code)) +
    geom_col() +
    geom_text(aes(label = paste0(round(100 * percent, 1), "%")), vjust = -.5, size=5)+
    xlab("Adverse Event Diagnosis") + 
    ylab("Count")+
    guides(fill=FALSE)+
    theme_bw()+
    theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20))
```

## Physical Restraints

```{r}
table(pwbe$Physical.Restraints, exclude = "ifany")
# No Yes 
#396  57 

pwbe %>% 
    group_by(Physical.Restraints) %>% 
    summarize(count = n()) %>% 
    mutate(percent = count/sum(count)) %>% 
    ggplot(aes(x=Physical.Restraints, y=count, fill=Physical.Restraints)) +
    geom_col() +
    geom_text(aes(label = paste0(round(100 * percent, 1), "%")), vjust = -.5, size=5)+
    xlab("Physical Restraints") + 
    ylab("Count")+
    guides(fill=FALSE)+
    theme_bw()+
    theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20))
```

# Percentage of patients with multiple yes in two or more columns

```{r}
vars <- c("Delirium.diagnosis", 
          "Antipsychotic.administration.while.on.Keppra", 
          "Positive.CAM.ICU.while.on.Keppra",
          "Adverse.event.diagnosis.code",
          "Physical.Restraints")

table(apply(pwbe[, which(names(pwbe) %in% vars)], 1, function(x){sum(x=="Yes")}))
#  1   2   3   4   5 
#246 127  54  23   3 

pwbe$multipleYes <- apply(pwbe[, which(names(pwbe) %in% vars)], 1, function(x){sum(x=="Yes")})
pwbe$multipleYes <- as.character(pwbe$multipleYes)

pwbe %>% 
    group_by(multipleYes) %>% 
    summarize(count = n()) %>% 
    mutate(percent = count/sum(count)) %>% 
    ggplot(aes(x=multipleYes, y=count, fill=multipleYes)) +
    geom_col() +
    geom_text(aes(label = paste0(round(100 * percent, 1), "%")), vjust = -.5, size=5)+
    scale_x_discrete(label=c("Only 1 Yes",
                              "2 Yes",
                              "3 Yes",
                              "4 Yes",
                              "5 Yes")) +
    xlab("Multiple Yes") + 
    ylab("Count")+
    guides(fill=FALSE)+
    theme_bw()+
    theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20))
```

## Patients with Antipsychotic administration while on Keppra; most frequently administered drug

```{r}
pwbe_antipsy <- pwbe[which(pwbe$Antipsychotic.administration.while.on.Keppra=="Yes"),]

table(pwbe_antipsy$Antipsychotic.received.while.on.Keppra, exclude = "ifany")
#  H  HO  HQ HQZ HRQ HZQ   O  OH  OQ OQH OQZ   Q  QH QHO QHZ  QO  QR QRH  QZ   R  RQ   Z  ZQ 
# 30   4  16   1   1   1   4   2   3   1   1  92  21   1   1   1   1   1   1   4   2   1   1 

pwbe_antipsy$H <- ifelse(grepl("H", pwbe_antipsy$Antipsychotic.received.while.on.Keppra), "Yes", "No")
pwbe_antipsy$O <- ifelse(grepl("O", pwbe_antipsy$Antipsychotic.received.while.on.Keppra), "Yes", "No")
pwbe_antipsy$Q <- ifelse(grepl("Q", pwbe_antipsy$Antipsychotic.received.while.on.Keppra), "Yes", "No")
pwbe_antipsy$Z <- ifelse(grepl("Z", pwbe_antipsy$Antipsychotic.received.while.on.Keppra), "Yes", "No")
pwbe_antipsy$R <- ifelse(grepl("R", pwbe_antipsy$Antipsychotic.received.while.on.Keppra), "Yes", "No")

apply(pwbe_antipsy[,c("H","O","Q","Z","R")], 2, table)

antipsychotic <- data.frame(Antipsychotic = c("H","O","Q","Z","R"),
                            Count = apply(pwbe_antipsy[,c("H","O","Q","Z","R")], 2, table)[2,])

ggplot(antipsychotic, aes(x=Antipsychotic, y=Count, fill=Antipsychotic)) +
    geom_col() +
    #geom_text(aes(label = paste0(round(100 * percent, 1), "%")), vjust = -.5, size=5)+
    xlab("Antipsychotic") + 
    ylab("Count")+
    guides(fill=FALSE)+
    theme_bw()+
    theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20))
```

# Time to event

## Time to Antipsyhotic Administration after first dose of Keppra (DAYS)

```{r}
pwbe$TTA_Status <- ifelse(pwbe$Antipsychotic.administration.while.on.Keppra=="Yes", 1, 0)
pwbe$TTA <- pwbe$Time.to.Antipsyhotic.Administration.after.first.dose.of.Keppra..DAYS.
pwbe$TTA[which(is.na(pwbe$TTA))] <- pwbe$Hospital.Length.of.Stay..DAYS.[which(is.na(pwbe$TTA))]

fittreat <- survfit(Surv(TTA, TTA_Status) ~ 1, data=pwbe)
survfit(Surv(TTA, TTA_Status) ~ 1, data=pwbe) #median=29.2
ggsurvplot(fittreat, 
            data = pwbe, 
            conf.int = TRUE, 
            censor=TRUE, 
            #size = 1.5,
            legend="none",
            xlab="Time to antipsychotic administration (days)",
            ylab="Probability of antipsychotic administration",
            risk.table = FALSE,
            fontsize = 5,
            font.x = 20,
            font.y = 20,
            ggtheme = theme_bw())
```

# Time to Positive CAM-ICU

```{r}
pwbe$PosCAM_Status <- ifelse(pwbe$Positive.CAM.ICU.while.on.Keppra=="Yes", 1, 0)
pwbe$PosCAM <- pwbe$Time.to.Positive.CAM.ICU.after.first.dose.of..Keppra..DAYS.
pwbe$PosCAM[which(is.na(pwbe$PosCAM))] <- pwbe$ICU.length.of.stay..DAYS.[which(is.na(pwbe$PosCAM))]

fittreat <- survfit(Surv(PosCAM, PosCAM_Status) ~ 1, data=pwbe)
survfit(Surv(PosCAM, PosCAM_Status) ~ 1, data=pwbe) #median=
ggsurvplot(fittreat, 
            data = pwbe, 
            conf.int = TRUE, 
            censor=TRUE, 
            #size = 1.5,
            legend="none",
            xlab="Time to positive CAM (days)",
            ylab="Probability of positive CAM",
            risk.table = FALSE,
            fontsize = 5,
            font.x = 20,
            font.y = 20,
            ggtheme = theme_bw())
```

